Paper 3: Housing and the capital share of income

### 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
prep OECD data

```{r}
library(tidyverse)
```

@@

files<- c("32_countries",
          "Canada",
          "US_alt")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
)

@@

@@

df <- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df<- df %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df <- df %>% 
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

@@

## Calculatiing derived variables

>>>>

```{r}
library(tidyverse)
library(readxl) # For handling xls source data files

@@

files<- c("32_countries",
          "Canada",
          "US_alt",
          "Japan")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx",
               "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"
)

@@

```{r}
df_list<- list()
```

@@

df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

@@

Data for Canada, Japan and an alternative (longer) series for the US.
```{r}
countries<- c("Canada", "US_alt", "Japan")

for (country in countries){
df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
                                                         sheet = 'DATA'))

# convert to long format and rename variables
df_list[[country]]<- df_list[[country]] %>%
  pivot_longer(cols = starts_with("_"), names_to = "year",
    names_prefix = "_", values_to ="value",
    names_transform = list(year=as.integer)) %>%
  rename(transaction = Variable) %>%
  mutate(country = country)

}

Tidying Canada data
```{r}
# Drop some NA values (- data for 2009 and 2010 is given as missing)
df_list[["Canada"]]<- df_list[["Canada"]] %>%
  filter(!is.na(value))

# For Canada we need to calculate some industry aggregates to be comparable with other countries
agg_industries<- list()

agg_industries[['D']]<- c(
  "15t16",
  "17t19",
  "20",
  "21t22",
  "23",
  "24",
  "25",
  "26",
  "27t28",
  "29",
  "30t33",
  "34t35",
  "36t37")

agg_industries[['G']]<- c(
  "50",
  "52")

agg_industries[['I']]<- c(
  "60t63",
  "64")

agg_industries[['K']]<- c(
  "70",
  "71t74")

for (agg_ind in names(agg_industries)){
  print(agg_ind)

  df_totals<- df_list[["Canada"]] %>%
  filter(code %in% agg_industries[[agg_ind]]) %>%
    group_by(transaction, year) %>%
    summarise(value = sum(value)) %>%
    ungroup() %>%
    mutate(code = agg_ind) %>%
    mutate(country = "Canada")

  df_list[["Canada"]]<- bind_rows(df_list[["Canada"]], df_totals)

}

```
Tidying Japan data
```{r}
# Drop NA values
df_list[["Japan"]]<- df_list[["Japan"]] %>% filter(is.na(value))

# Rename the capital compensation variable, to agree with other files
df_list[["Japan"]]<- df_list[["Japan"]] %>%
  mutate(transaction = replace(transaction,
                               transaction == "CAP_GFCF",
                               "CAP"))

Join the different dataframes together into a single dataframe.
```{r}
df<- bind_rows(
  df_list
)

Make a dictionary of industry codes and there descriptions from the combined data sources.
```{r}

df_industries<- df %>%
  select(code, desc) %>%
  drop_na() %>%
  unique()

# There are a couple of different spellings etc. from the different datasets. Here we just take select one (arbitrarily)
df_industries<- df_industries %>%
  distinct(code, .keep_all = TRUE)

# Add in the uniform industry code descriptions
df<- df %>%
  select(-desc)

df<- left_join(df, df_industries)

Some more tidying
```{r}
df<- df %>%
  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))

df<- df %>%
  pivot_wider(names_from = "transaction",
              values_from = "value")

# Any industry where VA = CAP but LAB is missing, set LAB = 0
  # This is generally only the case for sector Q, where VA=CAP=0.
df<- df %>%
  mutate(LAB = replace(LAB,
                       VA==CAP & is.na(LAB),
                       0))

# Any industry where VA = LAB but CAP is missing, set CAP = 0
  # This is only for sector P for US (NAICS data) and Korea (and Japan where P VA = 0)
df<- df %>%
  mutate(CAP = replace(CAP,
                       VA==LAB & is.na(CAP),
                       0))

# Drop any country-year-industry obs that do not have data for all 4 transactions
  # To-do check if COMP is needed, and calculate for US_alt and Japan if so)

df<- df %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )

Drop USA-NAICS data (in favour of of SIC data)
```{r}
df<- df %>%
  filter(country!="USA-NAICS")
```

Standardize country names
```{r}

# Mapping file
fp<- "PhD_pages/data/original/KLEMS/KLEMS_country_name_mapping.csv"

KLEMS_country_names<- read.csv(fp)

KLEMS_country_names<- KLEMS_country_names %>%
  rename(country = Country)

# aggregate shares
df<- merge(df, KLEMS_country_names)

df <- df %>%
  select(-country) %>%
  rename(country = Our.World.In.Data.Name)

## Complete dataset for aggregate industries

```{r}

# This list gives the codes of the main industries
main_industry_codes<- c("AtB",
                        "C",
                        "D",
                        "E",
                        "F",
                        "G",
                        "H",
                        "I",
                        "J",
                        "K",
                        "L",
                        "M",
                        "N",
                        "O",
                        "P",
                        "Q")

# filter for these codes sub-industries
industry_level_values<- df %>%
  filter(code %in% main_industry_codes)

# Store reported totals separately

reported_total_values<- df %>%
  filter(code == "TOT") %>%
  select(-code) %>%
  gather(transaction, reported_total, -c(country, year)) %>%
  arrange(country, year, transaction)

### 2022-10-31 PhD_pages/paper_pages/paper3.qmd
prep OECD data

>>>>

## Introduction

A substantial literature has documented a long-run shift in the distribution of national income between wages and returns to capital.

Much of the interest in this trend is due to it lying at the junction of two prominent topics within the social sciences: 1) the decline of labour bargaining power, and 2) the rise in income inequality.

(Explain (1) and connection to (2)).
